---
title: "Case Study 2 EDA only"
author: "Jackson Au & Chad Madding"
date: "November 25, 2018"
output:
 html_document:
   keep_md: yes
 pdf_document: default
 word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Install if the package doesn't exist 
#install.packages('DataExplorer') 
library(DataExplorer)
library(readr)
```

    MSDS 6306: Doing Data Science
    Case Study 02
    Due: Sunday, December 9th at 11:59pm. 

#### Description:
DDSAnalytics is an analytics company that specializes in talent management solutions for Fortune 1000 companies. Talent management is defined as the iterative process of developing and retaining employees. It may include workforce planning, employee training programs, identifying high-potential employees and reducing/preventing voluntary employee turnover (attrition). To gain a competitive edge over its competition, DDSAnalytics is planning to leverage data science for talent management. The executive leadership has identified predicting employee turnover as its first application of data science for talent management. Before the business green lights the project, they have tasked your data science team to conduct an analysis of existing employee data.

```{r read in data, echo=TRUE}
#read in data
training_attrition <- read.csv("CaseStudy2-data.csv", header=T,na.strings=c(""))

#data prep and cleaning
#check for NAs
sapply(training_attrition,function(x) sum(is.na(x)))

#Graph of missing data
plot_missing(training_attrition)

#we can drop ID, EmployeeCount, EmployeeNumber, Over18, StandardHours
training_attrition <- subset(training_attrition,select=c(2,3,4,5,6,7,8,9,12,13,14,15,16,17,18,19,20,21,22,24,25,26,27,29,30,31,32,33,34,35,36,37))

```

#### Goal:
Conduct exploratory data analysis (EDA) to determine factors that lead to attrition.

```{r exploratory data analysis, echo=TRUE}

#Variables
plot_str(training_attrition)

#Continuous Variables
plot_histogram(training_attrition)

plot_density(training_attrition)

#bivariate/multivariate analysis

#Correlation analysis
plot_correlation(training_attrition, type = 'continuous','Attrition')

#Categorical Variables-Barplots
plot_bar(training_attrition)
```

Looking at the first full model.
```{r first full model, echo=TRUE}
#checking out the first model
model <- glm(Attrition ~.,family=binomial(link='logit'),data=training_attrition)
#print out the model
summary(model)
```



Create a full EDA report

```{r full EDA report, echo=FALSE}
#Create a full EDA report
create_report(data=training_attrition, output_file = 'Case Study 2 EDA.html', y='Attrition')
```
